---
title: "Homework 3"
author: "Emirhan Bugday"
date: "November 19, 2018"
output:
  html_document:
    highlight: tango
    theme: simplex
    toc: yes
    toc_float:
      collapsed: yes
---

```{r, echo = F, warning = F, message=F}
library(data.table)
library("plot3D")
library(penalized)

# Set working directory and folder paths
setwd("G:\\My Drive\\Z.ME\\Boun\\IE 582\\HW-3")
load("Q1-Final.RData")
load("Q2-Final.RData")


```

## TASK 1

### Task 1 - A {.tabset .tabset-fade}

3D plots of the gestures are below (plot3D package is used to draw 3D plots of the instances, which represents 3D image in 2D).

#### Gesture 1
```{r echo= F}
i <- 11
instance <- data.table(Label = traindata_x[i,1], X_a = traindata_x[i,-1], Y_a = traindata_y[i,-1], Z_a = traindata_z[i,-1])

# Calculate Position from accerelation
instance[, X := cumsum(X_a)]
instance[, Y := cumsum(Y_a)]
instance[, Z := cumsum(Z_a)]

points3D(x = instance$X, y = instance$Y, z = instance$Z, pch=".", cex=7)

```

#### Gesture 2
```{r echo= F}
i <- 15
instance <- data.table(Label = traindata_x[i,1], X_a = traindata_x[i,-1], Y_a = traindata_y[i,-1], Z_a = traindata_z[i,-1])

# Calculate Position from accerelation
instance[, X := cumsum(X_a)]
instance[, Y := cumsum(Y_a)]
instance[, Z := cumsum(Z_a)]

points3D(x = instance$X, y = instance$Y, z = instance$Z, pch=".", cex=7)

```

#### Gesture 3
```{r echo= F}
i <- 4
instance <- data.table(Label = traindata_x[i,1], X_a = traindata_x[i,-1], Y_a = traindata_y[i,-1], Z_a = traindata_z[i,-1])

# Calculate Position from accerelation
instance[, X := cumsum(X_a)]
instance[, Y := cumsum(Y_a)]
instance[, Z := cumsum(Z_a)]

points3D(x = instance$X, y = instance$Y, z = instance$Z, pch=".", cex=7)

```

#### Gesture 4
```{r echo= F}
i <- 5
instance <- data.table(Label = traindata_x[i,1], X_a = traindata_x[i,-1], Y_a = traindata_y[i,-1], Z_a = traindata_z[i,-1])

# Calculate Position from accerelation
instance[, X := cumsum(X_a)]
instance[, Y := cumsum(Y_a)]
instance[, Z := cumsum(Z_a)]

points3D(x = instance$X, y = instance$Y, z = instance$Z, pch=".", cex=7)

```

#### Gesture 5
```{r echo= F}
i <- 2
instance <- data.table(Label = traindata_x[i,1], X_a = traindata_x[i,-1], Y_a = traindata_y[i,-1], Z_a = traindata_z[i,-1])

# Calculate Position from accerelation
instance[, X := cumsum(X_a)]
instance[, Y := cumsum(Y_a)]
instance[, Z := cumsum(Z_a)]

points3D(x = instance$X, y = instance$Y, z = instance$Z, pch=".", cex=7)

```

#### Gesture 6
```{r echo= F}
i <- 1
instance <- data.table(Label = traindata_x[i,1], X_a = traindata_x[i,-1], Y_a = traindata_y[i,-1], Z_a = traindata_z[i,-1])

# Calculate Position from accerelation
instance[, X := cumsum(X_a)]
instance[, Y := cumsum(Y_a)]
instance[, Z := cumsum(Z_a)]

points3D(x = instance$X, y = instance$Y, z = instance$Z, pch=".", cex=7)

```

#### Gesture 7
```{r echo= F}
i <- 7
instance <- data.table(Label = traindata_x[i,1], X_a = traindata_x[i,-1], Y_a = traindata_y[i,-1], Z_a = traindata_z[i,-1])

# Calculate Position from accerelation
instance[, X := cumsum(X_a)]
instance[, Y := cumsum(Y_a)]
instance[, Z := cumsum(Z_a)]

points3D(x = instance$X, y = instance$Y, z = instance$Z, pch=".", cex=7)

```

#### Gesture 8
```{r echo= F}
i <- 6
instance <- data.table(Label = traindata_x[i,1], X_a = traindata_x[i,-1], Y_a = traindata_y[i,-1], Z_a = traindata_z[i,-1])

# Calculate Position from accerelation
instance[, X := cumsum(X_a)]
instance[, Y := cumsum(Y_a)]
instance[, Z := cumsum(Z_a)]

points3D(x = instance$X, y = instance$Y, z = instance$Z, pch=".", cex=7)

```


### Task 1 - B

"Manhattan" & "Euclidean" distances along with k = 1, 3, 5, 7 is compared with 10 fold Cross Validation. 1 resampling is performed due to computational cost, however it would be more robust to select model parameters with more resampling. Since all predictors have the same unit, No pre-processing(normalization) is applied on data.

As a result, "manhattan" distance with k = 1 provided has the best performance. More detailed output of the experiment is brovided below.

```{r echo= F}
# Print Details
print(knnFit)

# Plot the Result
plot(knnFit)

```

### Task 1 - C
**For "Manhattan" Distance with k = 1:** 

Accuracy = 95.4%

Computation Time = 41.75 sec

& Confusion Matrix is as follows
```{r echo= F}
cm1
```

**For "Euclidean" Distance with k = 3:**

Accuracy = 94.4%

Computation Time = 4.88 sec

& Confusion Matrix is as follows
```{r echo= F}
cm2
```


## TASK 2

### Task 2 - A, B
Logistic regression model with fusel lasso penalty is trained on the training data, via "penalized" package.

lambda between 0 and 2 is tested for L1 penaltization on the differences of the coefficients while L1 penalization on the coefficients is set to 1 with 2 fold cross validation (2 fold is used due to computational costs)

Best lambda is found as 0.31 & 23 features are selected among 96. Summary of the result is provided below.

```{r echo= F}
fusedLasso
```

**Accuracy** of the model on the test data is **82%** and

**Confusion Matrix** is the following:
```{r echo= F}
T2_cm
```

Then, L1 penalization on the coefficients are optimized while lamda set for 0.31 for L1 penaltization on the differences of the coefficients. Results are provided below.
```{r echo= F}
fusedLasso_2
```

However, accuracy remained at **82%** and confusion matrix is stayed as the same.



### Task 2 - C, D
After the described time series transformation is applied, Logistic regression model with fusel lasso penalty is trained on the training data, via "penalized" package.

lambda between 0 and 2 is tested for L1 penaltization on the differences of the coefficients while L1 penalization on the coefficients is set to 1 with 2 fold cross validation (2 fold is used due to computational costs)

Best lambda is found as 0.19 & 18 features are selected among 95. Summary of the result is provided below.

```{r echo= F}
fusedLasso2
```

**Accuracy** of the model on the test data is **84%** and

**Confusion Matrix** is the following:
```{r echo= F}
T2_cm_2
```

Then, L1 penalization on the coefficients are optimized while lamda set for 0.19 for L1 penaltization on the differences of the coefficients. Results are provided below.
```{r echo= F}
fusedLasso2_2
```

As a result Accuracy is increased to **85%** and confusion matrix become as follows:
```{r echo= F}
T2_cm2_2
```

However, the total number of selected features are increased to **33**. Thus, we can prefer lambda1 = 1 with **84%** accuracy in order to deal with a smaller feature dimension.

To conclude, in the first setting we observe that 82% accuracy can be obtained 23/96 features and selecting addtional features among the originally provided ones doesn't contribut. On the other hand, with the described transformation- the second setting- slightly better accuracy can be achieved with even less features, 18/96.


## APPENDIX
Source code of Task 1 and Task 2

### TASK 1
```{r , eval = F, echo = T}
# Clear connections, environment
closeAllConnections()
rm(list = ls())
gc()

# Load Neccesary Packages
library(data.table)
library(caret)
library(tictoc)
library(kknn)

# Load Train Data
fname='G:\\My Drive\\Z.ME\\Boun\\IE 582\\HW-3\\UWave\\uWaveGestureLibrary_X_TRAIN' # data path
traindata_x <- as.matrix(read.table(fname))  # read data into a matrix named traindata

fname='G:\\My Drive\\Z.ME\\Boun\\IE 582\\HW-3\\UWave\\uWaveGestureLibrary_Y_TRAIN' # data path
traindata_y <- as.matrix(read.table(fname))  # read data into a matrix named traindata

fname='G:\\My Drive\\Z.ME\\Boun\\IE 582\\HW-3\\UWave\\uWaveGestureLibrary_Z_TRAIN' # data path
traindata_z <- as.matrix(read.table(fname))  # read data into a matrix named traindata

# Cbind Train Data and Change Column Names
traindata <- cbind(traindata_x, traindata_y[,-1])
traindata <- cbind(traindata, traindata_z[,-1])
traindata<- as.data.table(traindata)
names(traindata) <- as.character(c("gesture", 1:(dim(traindata)[2] - 1)))
# names(traindata) <- as.character(c("gesture", paste0("V", 1:(dim(traindata)[2] - 1))))
str(traindata)

# Load Test Data
fname='G:\\My Drive\\Z.ME\\Boun\\IE 582\\HW-3\\UWave\\uWaveGestureLibrary_X_TEST' # data path
testdata_x <- as.matrix(read.table(fname))  # read data into a matrix named traindata

fname='G:\\My Drive\\Z.ME\\Boun\\IE 582\\HW-3\\UWave\\uWaveGestureLibrary_Y_TEST' # data path
testdata_y <- as.matrix(read.table(fname))  # read data into a matrix named traindata

fname='G:\\My Drive\\Z.ME\\Boun\\IE 582\\HW-3\\UWave\\uWaveGestureLibrary_Z_TEST' # data path
testdata_z <- as.matrix(read.table(fname))  # read data into a matrix named traindata

testdata <- cbind(testdata_x, testdata_y[,-1])
testdata <- cbind(testdata, testdata_z[,-1])
testdata<- as.data.table(testdata)
names(testdata) <- as.character(c("gesture", 1:(dim(testdata)[2] - 1)))
# names(testdata) <- as.character(c("gesture", paste0("V", 1:(dim(testdata)[2] - 1))))
str(testdata)


#1-A
i <- 1
instance <- data.table(Label = traindata_x[i,1], X_a = traindata_x[i,-1], Y_a = traindata_y[i,-1], Z_a = traindata_z[i,-1])

# Calculate Position from accerelation
instance[, X := cumsum(X_a)]
instance[, Y := cumsum(Y_a)]
instance[, Z := cumsum(Z_a)]

par(mfrow=c(1,1))

library("plot3D")
points3D(x = instance$X, y = instance$Y, z = instance$Z, pch=".", cex=7)

# library(rgl)
# plot3d(x = instance$X, y = instance$Y, z = instance$Z, pch=".", cex=7)
# plot3d(x = instance$X, y = instance$Y, z = instance$time, pch=".", cex=7)


#1-B
# For Classification
traindata[, gesture := as.factor(gesture)]
testdata[, gesture := as.factor(gesture)]

# Create 10 fold CV data set
fitControl <- trainControl(## 10-fold CV
  method = "repeatedcv",
  number = 10,
  ## repeated ten times
  repeats = 1)

# Create Parameter grid to evaluate throught cross validation
# Selected Distances: 1 - Manhattan, 2 - Euclidean
# Tries k values: 1,3,5
# kernel = "rectangular" corresponds standard unweighted knn (it is required for the package)
knnGrid <-  expand.grid(kmax = c(1,3,5,8), distance = c(1,2), kernel = "rectangular")

# Optimize parameters using the train data set
knnFit <- train(gesture ~ ., data = traindata, 
                method = "kknn", 
                trControl = fitControl,
                tuneGrid = knnGrid,
                verbose = TRUE)

saveRDS(knnFit, "G:\\My Drive\\Z.ME\\Boun\\IE 582\\HW-3\\KnnFitQ1Final.Rds")

# Plot the Result
plot(knnFit)

# Print Details
print(knnFit)


# 1-C

tic("manhattan")
prediction1 <-  kknn(gesture ~ ., 
                     traindata, 
                     testdata, 
                     k = 1, 
                     distance = 1, 
                     kernel = "rectangular",
                     # na.action = na.omit(),
                     scale = FALSE)
toc()

result1 <- cbind(data.table(prediction = prediction1$fitted.values), data.table(actual = testdata$gesture))
cm1 <- table(result1)
accu1 <- sum(diag(cm1))/sum(cm1)

tic("euclidean")
prediction2 <-  kknn(gesture ~ ., 
                     traindata, 
                     testdata, 
                     k = 3, 
                     distance = 2, 
                     kernel = "rectangular",
                     # na.action = na.omit(),
                     scale = FALSE)
toc()

result2 <- cbind(data.table(prediction = prediction2$fitted.values), data.table(actual = testdata$gesture))
cm2 <- table(result2)
accu2 <- sum(diag(cm2))/sum(cm2)


save.image(file = "G:\\My Drive\\Z.ME\\Boun\\IE 582\\HW-3\\Q1-Final.RData")


```

### TASK 2
```{r , eval = F, echo = T}
# Clear connections, environment
closeAllConnections()
rm(list = ls())
gc()

library(penalized)
library(data.table)

# Load Train Data
fname='G:\\My Drive\\Z.ME\\Boun\\IE 582\\HW-3\\NN_Classification\\ecgTRAIN' # data path
traindata <- as.data.table(read.table(fname))  # read data into a matrix named traindata

# Train Class Variable
trainclass = traindata[, 1] # takes -1 and 1
# Convert class variable to 0-1
trainclass <- (trainclass + 1) / 2
str(trainclass) 
# Train Features
traindata = traindata[, 2:ncol(traindata)]

# Load Test Data
fname='G:\\My Drive\\Z.ME\\Boun\\IE 582\\HW-3\\NN_Classification\\ecgTEST' # data path
testdata <- as.matrix(read.table(fname))  # read data into a matrix named traindata

# Test Class Variable
testclass = testdata[, 1]
# Convert class variable to 0-1
testclass <- (testclass + 1) / 2
# Test Features
testdata=testdata[, 2:ncol(testdata)]


#4-A,B
fusedLasso <- optL2(response = as.matrix(trainclass), 
                    penalized = traindata, 
                    unpenalized = ~ 0, 
                    data = traindata,
                    fusedl = T, 
                    standardize = T, 
                    lambda1 = 1, 
                    minlambda2 = 0, 
                    maxlambda2 = 2,
                    fold = 2,
                    model = "logistic")


# pen <- penalized(response = as.matrix(trainclass),
#                  penalized = traindata,
#                  unpenalized = ~ 0,
#                  data = traindata,
#                  fusedl = T,
#                  standardize = F,
#                  lambda1 = 1,
#                  lambda2 = fusedLasso$lambda,
#                  model = "logistic",
#                  trace = T)

coef(fusedLasso$fullfit)

T2_prediction <- predict(object =  fusedLasso$fullfit, as.matrix(testdata))

T2_result <- cbind(data.table(prediction = round(T2_prediction, digit = 0)), data.table(actual = testclass))
T2_cm <- table(T2_result)
T2_accu <- sum(diag(T2_cm))/sum(T2_cm)


fusedLasso_2 <- optL1(response = as.matrix(trainclass), 
                    penalized = traindata, 
                    unpenalized = ~ 0, 
                    data = traindata,
                    fusedl = T, 
                    standardize = T, 
                    lambda2 = fusedLasso$lambda, 
                    minlambda1 = 0, 
                    maxlambda1 = 2,
                    fold = 2,
                    model = "logistic")

coef(fusedLasso_2$fullfit)

T2_prediction_2 <- predict(object =  fusedLasso_2$fullfit, as.matrix(testdata))

T2_result_2 <- cbind(data.table(prediction = round(T2_prediction_2, digit = 0)), data.table(actual = testclass))
T2_cm_2 <- table(T2_result_2)
T2_accu_2 <- sum(diag(T2_cm_2))/sum(T2_cm_2)


# 4-C,D

# Compute diff variables for train data
traindata_2 <- as.data.table(traindata)
traindata_2[,instance := c(1:.N)]
traindata_2_2 <- melt(traindata_2,id.vars = "instance")
traindata_2_2 <- traindata_2_2[order(instance, variable)]
traindata_2_2[, value2 := value - shift(value, n = 1, type = "lag"), instance]
traindata_2 <- dcast(traindata_2_2[!is.na(value2), .(instance, variable, value2)], formula = instance  ~ variable, value.var = "value2")
traindata_2[, instance := NULL]

summary(traindata_2)

# Compute diff variables for test data
testdata_2 <- as.data.table(testdata)
testdata_2[,instance := c(1:.N)]
testdata_2_2 <- melt(testdata_2,id.vars = "instance")
testdata_2_2 <- testdata_2_2[order(instance, variable)]
testdata_2_2[, value2 := value - shift(value, n = 1, type = "lag"), instance]
testdata_2 <- dcast(testdata_2_2[!is.na(value2), .(instance, variable, value2)], formula = instance  ~ variable, value.var = "value2")
testdata_2[, instance := NULL]

summary(testdata_2)


fusedLasso2 <- optL2(response = as.matrix(trainclass), 
                    penalized = traindata_2, 
                    unpenalized = ~ 0, 
                    data = traindata_2,
                    fusedl = T, 
                    standardize = T, 
                    lambda1 = 1, 
                    minlambda2 = 0, 
                    maxlambda2 = 2,
                    fold = 2,
                    model = "logistic")



coef(fusedLasso2$fullfit)

T2_prediction2 <- predict(object = fusedLasso2$fullfit, as.matrix(testdata_2))

T2_result2 <- cbind(data.table(prediction = round(T2_prediction2, digit = 0)), data.table(actual = testclass))
T2_cm2 <- table(T2_result2)
T2_accu2 <- sum(diag(T2_cm2))/sum(T2_cm2)

fusedLasso2$lambda



fusedLasso2_2 <- optL1(response = as.matrix(trainclass), 
                       penalized = traindata_2, 
                       unpenalized = ~ 0, 
                       data = traindata_2,
                       fusedl = T, 
                       standardize = T, 
                       lambda2 = fusedLasso2$lambda, 
                       minlambda1 = 0, 
                       maxlambda1 = 2,
                       fold = 2,
                       model = "logistic")

T2_prediction2_2 <- predict(object = fusedLasso2_2$fullfit, as.matrix(testdata_2))

T2_result2_2 <- cbind(data.table(prediction = round(T2_prediction2_2, digit = 0)), data.table(actual = testclass))
T2_cm2_2 <- table(T2_result2_2)
T2_accu2_2 <- sum(diag(T2_cm2_2))/sum(T2_cm2_2)

save.image(file = "G:\\My Drive\\Z.ME\\Boun\\IE 582\\HW-3\\Q2-Final.RData")


```
